---
title: "R Notebook"
output: html_notebook
---
```{r}
library("ggplot2")
library("dplyr")
library("ggsci")
library("ggpubr")
library("forcats")
accident_data <- read.csv("C:/Users/ranja/Downloads/Regulatory Affairs of Road Accident Data 2020 India.csv")
head(accident_data)
colnames(accident_data)
```
```{r}
accident_data["Outcome.of.Incident"]%>%unique()
accident_data["Cause.category"]%>%unique()
accident_data["Cause.Subcategory"]%>%unique()
```
```{r}
persons_killed <- accident_data%>%
  filter(Outcome.of.Incident == "Persons Killed")
a <- sum(persons_killed$Count, na.rm = TRUE)
persons_injured <- accident_data%>%
  filter(Outcome.of.Incident == "Greviously Injured")
b <- sum(persons_injured$Count, na.rm = TRUE)
minor_injuries <- accident_data%>%
  filter(Outcome.of.Incident == "Minor Injury")
c <- sum(minor_injuries$Count, na.rm = TRUE)
affected_people <- data.frame(Total.injured = b,
                              Total.killed = a,
                              Total.Minor.Injuries = c)
affected_people
```
```{r}
accident_data <- na.omit(accident_data)
```

```{r}
plot <- accident_data%>%
   filter(Outcome.of.Incident == "Persons Killed")%>%
  group_by(Cause.category)%>%
  summarize(Total.Death = sum(Count))%>%
  mutate(Causes = fct_reorder(Cause.category, Total.Death))
```
```{r}
bar_plot <- plot%>%ggplot(aes(x = Causes, y = Total.Death, fill = Cause.category)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.6) +
  scale_fill_npg() +
  labs(
    title = "Total Deaths by Cause",
    x = "Cause Subcategory",
    y = "Total Death",
    fill = "Cause Subcategory"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text = element_text(size = 6),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
print(bar_plot)

  
```
```{r}
cities_comparison <- accident_data%>%
  filter(Outcome.of.Incident == "Persons Killed")%>%
  group_by(Million.Plus.Cities)%>%
  summarise(Fatalities = sum(Count))%>%
  select(Cities = Million.Plus.Cities, Fatalities)
print(cities_comparison)
```
```{r}
plot <- cities_comparison%>%
  ggplot(
    aes(reorder(Cities, Fatalities), Fatalities)
  )+
  geom_bar(
    fill = "lightgreen", colour = "black",
           stat = "identity",
           width = 0.9)+
  geom_text(
    aes(label = Fatalities),
    cex = 2.5,
    size = 0.9,
    angle = 90
  )+
  labs(
    title = "Fatalities by City",
    x = "Cities",
    y = "Total Deaths"
  )+
  theme_minimal()+
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7, colour = "forestgreen"),
    axis.text.y = element_text(hjust = 1, size = 12, colour = "maroon"),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
print(plot)
```
So, from this plot we got the conclusions that :
1. Delhi is the city here most people got killed by accidents in 2020.
2. Srinagar is the city where fatalities caused by accidents are lowest.


```{r}
injuries_by_cities <- accident_data%>%
  filter(Outcome.of.Incident == "Greviously Injured")%>%
  group_by(Million.Plus.Cities)%>%
  summarise(Serious.Injuries = sum(Count))%>%
  select(Cities = Million.Plus.Cities, Serious.Injuries)
print(injuries_by_cities)
```


```{r}
plot2 <- injuries_by_cities%>%
  ggplot(
    aes(reorder(Cities, Serious.Injuries), Serious.Injuries)
  )+
  geom_bar(
    fill = "cyan", colour = "black",
           stat = "identity",
           width = 0.9)+
  geom_text(
    aes(label = Serious.Injuries),
    size = 2,
    angle = 90
  )+
  labs(
    title = "Comparison of Cities for Accident With Serious Injuries",
    x = "Cities",
    y = "Number of People Injured"
  )+
  theme_minimal()+
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 7, colour = "forestgreen"),
    axis.text.y = element_text(hjust = 1, size = 12, colour = "maroon"),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
print(plot2)
```
From this plot we see that :
1. Most people that got seriously injured from road accidents in 2020 are from Bengaluru.
2. Srinagar and Thiruchirapally did a great job by achieving zero serious injuries in road accidents in 2020.


```{r}
plot_func <- function(cause, outcome) {
  accident_data%>%
    filter(Cause.category == cause, Outcome.of.Incident == outcome)%>%
    group_by(Cause.Subcategory)%>%
    summarise(Total.outcome = sum(Count))%>%
    ggplot(aes(reorder(Cause.Subcategory, Total.outcome), Total.outcome, fill = Cause.Subcategory))+
    geom_bar(stat = "identity", color = "black", width = 0.9)+
    geom_text(aes(label = Total.outcome), vjust = -0.5, size = 3)+
    scale_fill_frontiers()+
    labs(title = paste(cause, "-", outcome))+
    theme_light()+
    theme(
       plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, colour = "forestgreen"),
    axis.text.y = element_text(hjust = 1, size = 12, colour = "maroon"),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
    )
}
combinations <- expand.grid(
  cause = unique(accident_data$Cause.category),
  outcome = unique(accident_data$Outcome.of.Incident)
)
plots <- lapply(1:nrow(combinations), function(i) {
  plot_func(
    cause = combinations$cause[i],
    outcome = combinations$outcome[i]
  )
})
print(plots)
  
```

These plots give us the following conclusions:
1. Most accidents due to traffic control are caused by uncontrolled traffic.
2. Mostly people with two wheelers get seriously injured.
3. Most accidents in staggered junctions and other undefined junctions causes just minor injuries.


```{r}
output_dir <- "plots"
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}


for (i in seq_along(plots)) {
  ggsave(
    filename = paste0(output_dir, "/plot_", i, ".png"), 
    plot = plots[[i]],  
    width = 6,          
    height = 4          
  )
}

```

```{r}
summary_func <- function(outcome) {
  accident_data %>%
    filter(Outcome.of.Incident == outcome) %>%
    group_by(Cause.category, Cause.Subcategory) %>%
    summarise(Total.outcome = sum(Count), .groups = "drop") %>%
    arrange(desc(Total.outcome)) %>%
    slice_head(n = 1) 
}

outcomes <- c("Persons Killed", "Greviously Injured", "Minor Injury")
top_causes <- lapply(outcomes, summary_func)
result <- bind_rows(
  lapply(seq_along(outcomes), function(i) {
    top_causes[[i]] %>%
      mutate(Outcome = outcomes[i])
  })
)

print(result)
```
This table shows that most people affected due to accidents are caused by Sunny/clear weather and straight roads which don't make any sense if there is not overspeeding involved.

